---
# Copyright 2017 Verily Life Sciences Inc.
#
# Use of this source code is governed by a BSD-style
# license that can be found in the LICENSE file.

title: "Quality Control"
output: html_document
params:
  PROJECT_ID: "YOUR_PROJECT_ID"
  DATASET_DESCRIPTION: "Brief description of the single-cell dataset."
  # This table must exist.
  RAW_DATA_TABLE: "PROJECT_ID_THE_DATA_IS_IN.DATASET_NAME.TABLE_NAME"
  # These tables will be created.
  CELL_METRICS_TABLE: "YOUR_DATASET_NAME.TABLE_NAME"
  PASSING_CELLS_TABLE: "YOUR_DATASET_NAME.TABLE_NAME"
  GENE_METRICS_TABLE: "YOUR_DATASET_NAME.TABLE_NAME"
  PASSING_GENES_TABLE: "YOUR_DATASET_NAME.TABLE_NAME"

# This RMarkdown is a parameterized report.  See
# http://rmarkdown.rstudio.com/developer_parameterized_reports.html
# for more detail.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Whether or not to cache chunk results for faster rendering when tweaking narrative
# and plots.  To avoid confusion, set this to FALSE and rely on BigQuery's server-side
# caching when actively modifying SQL.
knitr::opts_chunk$set(cache=TRUE)
```

This report performs quality control on dataset: `r params$DATASET_DESCRIPTION`

```{r}
library(dplyr)
library(ggplot2)
library(scales)
# Use the copy from GitHub, not CRAN, which has support for standard SQL.
# install.packages("devtools")
# devtools::install_github("rstats-db/bigrquery")
library(bigrquery)
```

There are many ways to facilitate templated queries.  Here we use Python via [reticulate](https://github.com/rstudio/reticulate) and [Jinja2](http://jinja.pocoo.org/docs/2.9/). As another example, see [this R approach](https://github.com/googlegenomics/codelabs/blob/3a0a1b754e78bc159a9c6deb604a60344034cc2a/R/PlatinumGenomes-QC/rHelpers/setup.R).

```{r helper, cache=FALSE}
library(reticulate)
jinja = import("jinja2")
py = import_builtins()

perform_bqquery = function(sql_path, ...) {
  sql = jinja$Template(py$open(sql_path, "r")$read())$render(params)
  cat(sql)
  query_exec(sql, useLegacySql = FALSE, project=params$PROJECT_ID, ...)
}

```

## Compute raw data counts

```{r raw_counts}
perform_bqquery(sql_path = "raw_data_counts.sql")
```

## Compute cell metrics

```{r warning=FALSE}
cell_metrics = perform_bqquery(
  sql_path = "cell_metrics.sql",
  destination_table = params$CELL_METRICS_TABLE)

```

## Visualize cell metrics

```{r}
qplot(alltrans, data=cell_metrics, geom="histogram")
```

```{r}
qplot(mttrans, data=cell_metrics, geom="histogram")
```

```{r}
qplot(gene_cnt, data=cell_metrics, geom="histogram")
```

## Choose quality control thresholds and compute the cell subset

```{r passing_cells, message=FALSE}
perform_bqquery(sql_path = "passing_cells.sql",
                destination_table = params$PASSING_CELLS_TABLE,
                page_size = 10,
                max_pages = 1)
```

## Compute gene metrics

```{r gene_metrics, warning=FALSE}
gene_metrics = perform_bqquery(
  sql_path = "gene_metrics.sql",
  destination_table = params$GENE_METRICS_TABLE)

```

## Visualize gene metrics

```{r}
qplot(alltrans, data=gene_metrics, geom="histogram")
```

```{r}
qplot(cell_cnt, data=gene_metrics, geom="histogram")
```

## Choose quality control thresholds and compute the gene subset

```{r passing_genes, message=FALSE}
perform_bqquery(sql_path = "passing_genes.sql",
                destination_table = params$PASSING_GENES_TABLE,
                page_size = 10,
                max_pages = 1)
```

## Compute filtered data counts

```{r passing_counts, message=FALSE}
perform_bqquery(sql_path = "passing_data_counts.sql")
```
